Estimating the Financial Impact of Wildfire Smoke on US National parks

### 2021-05-11 MarkdownDocuments/Manuscript.Rmd
got the equations rendered

---
title: Estimating the Financial Impact of Wildfire Smoke on US National parks
author:
- affiliation: Boise State University - Human Environment Systems
  name: Matt Clark
- affiliation: University of Michigan
  name: Alexander Killion
- affiliation: Boise State University
  name: Matthew A. Williamson
- affiliation: Boise State University
  name: Vicken?
- affiliation: Boise State University
  name: Laura? (undergrad who will continue working over summer)
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
  word_document: default
fontfamily: mathpazo
fontsize: 11pt
geometry: margin=1in
keywords: Wildfire, Smoke, Recreation, Park Science
bibliography: ../References.bib
thanks: If we want to thank anyone? Maybe Moji?.
abstract: The United States national park system serves as a global model for how natural resource protection can provide numerous social and ecological services while generating economic returns on investment. In recent years parks have increasingly relied on visitation fees to fund conservation projects as well as basic infrastructure. In contrast with congressional funding, visitation is subject to fluctuation given climatic and environmental variability. While many environmental variables are relatively predictable and stable over time, wildfire smoke events have increased dramatically via climate change, particularly in the American West. In this project, we assess how recreational visitation to national parks has systematically varied with wildfire smoke. Over the last 38 years, we estimate that wildfire smoke has led to a 12% overall visitation decrease to National Parks in the Pacific West and Inter-Mountain Regions. When accounting for visitation fees, this loss may have cost the National Park Service as much as $580 million in total. Wildfire frequency and severity are expected to increase considerably in these regions over the next 50 years. Assuming current trends in smoke, park visitation, and park fees continue, we expect to see significant financial losses to the National Park Service via wildfire smoke in the coming years.     
---
# 1. Introduction

>>>>

---
title: Estimating the Financial Impact of Wildfire Smoke on US National parks
author:
- affiliation: Boise State University - Human Environment Systems
  name: Matt Clark
- affiliation: University of Michigan
  name: Alexander Killion
- affiliation: Boise State University
  name: Matthew A. Williamson
- affiliation: Boise State University
  name: Vicken?
- affiliation: Boise State University
  name: Laura? (undergrad who will continue working over summer)
date: "`r format(Sys.time(), '%B %d, %Y')`"
  output:
  html_document:
    df_print: paged
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
  word_document: default
fontfamily: mathpazo
fontsize: 11pt
template: ~/SmokeProject/MarkdownDocuments/MillerTemplate.tex

geometry: margin=1in
keywords: Wildfire, Smoke, Recreation, Park Science
bibliography: ../References.bib
thanks: If we want to thank anyone? Maybe Moji?.
abstract: The United States national park system serves as a global model for how natural resource protection can provide numerous social and ecological services while generating economic returns on investment. In recent years parks have increasingly relied on visitation fees to fund conservation projects as well as basic infrastructure. In contrast with congressional funding, visitation is subject to fluctuation given climatic and environmental variability. While many environmental variables are relatively predictable and stable over time, wildfire smoke events have increased dramatically via climate change, particularly in the American West. In this project, we assess how recreational visitation to national parks has systematically varied with wildfire smoke. Over the last 38 years, we estimate that wildfire smoke has led to a 12% overall visitation decrease to National Parks in the Pacific West and Inter-Mountain Regions. When accounting for visitation fees, this loss may have cost the National Park Service as much as $580 million in total. Wildfire frequency and severity are expected to increase considerably in these regions over the next 50 years. Assuming current trends in smoke, park visitation, and park fees continue, we expect to see significant financial losses to the National Park Service via wildfire smoke in the coming years.     
---
# 1. Introduction
